---
title: "Difference in Differences Estimators Lab"
author: "Zahid Asghar<br><br> School of Economics, QAU, Islamabad"
date: today
format: 
  html:
    self-contained: false
    slide-number: c/t
    width: 1600
    height: 900
    logo: "https://www.rstudio.com/wp-content/uploads/2018/10/RStudio-Logo-Flat.png"
    footer: "[Zahid Asghar](https://zahidasghar.com/econometrics/Panel_reg)"
    theme: ["simple", "styles.scss"]
    echo: true
    chalkboard: true
    multiplex: true
    code-link: true
    title-slide-attributes:
      data-background-color: "#447099"
editor: source
---

```{r include=FALSE}
library(tidyverse)
library(httr)
clrs <- MetBrewer::met.brewer(name = "Java")
clrs_lt <- colorspace::lighten(clrs, 0.9)
knitr::opts_chunk$set(fig.retina = 3, collapse = TRUE)
options(digits = 3, width = 75)
```

Purpose of this lab is to understand the concept of DiD method and its usefulness in R. For more details, one may visit [Tyler Rasom](https://github.com/tyleransom/EconometricsLabs/blob/master/lab17.Rmd)

## Upload relevant librarires

Open a quarto document or R script in `RStudio`, which is now named as `Posit` and run these libraries. If some `library()` does not work, make sure you have uploaded the relevant `package`.

```{r}
# Add names of group members HERE
library(tidyverse)
library(wooldridge)
library(broom)
library(magrittr)
library(stargazer)
library(clubSandwich)
```

### Load the data

We are using data analysed by @meyer_al1995 and is available from `wooldridge` package and filter only to those living in Knetucky:

```{r}
df <- as_tibble(injury)
df %<>% filter(ky==1)
```

### Policy setting

In 1980, Kentucky raised its cap on weekly earnings that were covered by worker's compensation.[^1] The outcome variable is `ldurat`, which is the log duration (in weeks) of worker's compensation benefits. The policy was such that the cap increase did not affect low-earnings workers, but did affect high-earnings workers. Thus, low-earnings workers serve as the control group, while high-earnings workers serve as the treatment group.

[^1]: Worker's compensation is "a form of insurance providing wage replacement and medical benefits to employees injured in the course of employment in exchange for mandatory relinquishment of the employee's right to sue their employer for ... negligence." ([Wikipedia](https://en.wikipedia.org/wiki/Workers%27_compensation))

We are interested to know if the policy caused workers to spend more time out of work. If benefits are not generous enough, then workers may sue the company for on-the-job injuries. On the other hand, benefits that are too generous may induce workers to be more reckless on the job, or to claim that off-the-job injuries were incurred while at work.

[zahid asghar](zahidasghar.com)

### Summary statistics

In difference-in-differences settings, it is often helpful to see if there even is a policy effect. Let's compute the difference-in-differences (with no regression) based on the pre- and post-means for the treatment and control groups

```{r}
df %>% group_by(afchnge,highearn) %>% summarize(mean.ldurat = mean(ldurat))
```

1.  What is the difference in the differences? What is its interpretation? (Hint: recall that the dependent variable is logged)

## Difference-in-differences regression

The basic regression model to analyze the policy's impact is $log(durat) = \beta_0 + \delta_0 afchnge + \beta_1 highearn + \delta_1 afchnge \cdot highearn + u$

Estimate this model, calling it `est.did`. I suppress the code here, since this should be old hat by now. (Hint: recall that, for two dummy variables `x1` and `x2`, putting `x1*x2` in the `lm()` formula will include in the formula each dummy and the interaction of the two.)

```{r }
est.did <- lm(ldurat ~ afchnge*highearn, data=df)
```

2.  Verify that your estimated $\hat{\delta}_1$ is the same as your answer in (1). Is the effect significant?

## Include additional regressors

We might want to control for other aspects of our workers. For example, perhaps claims made by construction or manufacturing workers tend to have longer duration than claims made workers in other industries. Or maybe those claiming back injuries tend to have lonter claims than those claiming head injuries. One may also want to control for worker demographics such as gender, marital status, and age.

Estimate an expanded version of the basic regression model, where the following additional variables are included:

-   `male`
-   `married`
-   Quadratic in `age`
-   `hosp` (1 = hospitalized)
-   `indust` (1 = manuf, 2 = construc, 3 = other)
-   `injtype` (1-8; categories for different types of injury)
-   `lprewage` (log of wage prior to filing a claim)

**Be sure to format `indust` and `injtyp` as factors before proceeding.** Call your model `est.did.x`.

```{r }

df %<>% mutate(indust=as.factor(indust), injtype=as.factor(injtype))
est.did.x <- lm(ldurat ~ afchnge*highearn + male + married + age + I(age^2) + injtype + lprewage + indust, data=df)
```

We now view results from `modelsummary`

```{r}
library(modelsummary)
modelsummary(list(est.did,est.did.x),estimate = "{estimate}{stars}",
             gof_omit = ".*")
```

3.  How different is your estimate of $\hat{\delta}_1$ from that in (2)?

4.  Do you see any interesting patterns among the $x$'s in predicting log duration of benefits?

# References
